A Stopped Negative Binomial Distribution

### 2016-06-15 stat-and-prob-letters/snb-paper.tex
Yeah.

@@

\begin{figure}[t!]
\includegraphics[width=\textwidth]{KanePlot.pdf}
\caption{
A hypothetical realization of the trial.
}
\label{fig:kane_viz}
\end{figure}

@@

In this work we derive the distribution of the number of enrollees needed
to observe either $s$ successes or $t$ failures. We refer to this distribution
as the Stopped Negative Binomial (SNB). Some of its characteristics are
summarized in Tab.~\ref{tab:snb}.
The rest of this paper derives these results
and explores properties of the distribution.
%The next section introduces our notation and basic results
%including the density of the distribution along with a description of
%its relation to other distributions. 
Section 2 derives the distribution function
based on a defined Bernoulli process and gives some basic properties.
Section 3 shows how the distribution is related to other standard
distributions and connects the SNB tail probability to the Binomial tail 
probability.
Section 4 derives the moment generating function.
Section 5 derives the compound distribution using a Beta distribution
for $p$.

@@

\begin{figure}[t!]
\centering
\subfloat[The compound distribution's expected sample size.]{\includegraphics[width=0.5\textwidth]{bayesian-sample-expectation.pdf}}
\hfill
\subfloat[The compound distribution's variance.]{\includegraphics[width=0.5\textwidth]{bayesian-sample-variance.pdf}}
\caption{
The expected size and variance of the compound SNB with $s=2$, $t=12$
with $\mathbb{E}\ p = 0.045$ varying the uncertainty 
in the distribution of $p$. A larger value of $c$
corresponds to a Beta with smaller variance.
}
\label{fig:bayesian-sample-size}
\end{figure}

>>>>

\begin{figure}[bp!]
\includegraphics[width=\textwidth]{KanePlot.pdf}
\caption{
A hypothetical realization of the trial.
}
\label{fig:kane_viz}
\end{figure}

@@

\begin{figure}[bp!]
\centering
\includegraphics[width=\textwidth]{snb-first-plot.pdf}
\caption{
The probability mass function for a trial where $p$ 
is 0.045 and where the trial stops
after two patients respond or 11 patients do not respond.
}
\label{fig:first-plot}
\end{figure}

More generally the distribution for number of trial enrollments is shown in 
Fig.~\ref{fig:first-plot}. The probability mass is concentrated at the 
$11^{\text{th}}$ and $12^{\text{th}}$ step since $p$ 
is small and for most realizations reach 11 
non-responders before reaching 2 responders.

\begin{figure}[bp!]
\centering
\subfloat[Expected value]{\includegraphics[width=0.5\textwidth]{dist-mean.pdf}}
\hfill
\subfloat[Variance]{\includegraphics[width=0.5\textwidth]{dist-var.pdf}}
\caption{
The mean and variance for a trial that stops after two patients respond
or 11 patients do not respond when $p$ varies between zero and one.
}
\label{fig:exp-and-var}
\end{figure}

Fig.~\ref{fig:exp-and-var} shows the expected value and variance for the
number of trial enrollments varying $p$ between zero an 1. When $p$ is
small the probability of an individual response is small, the treatment
is more likely to fail, and the number of enrollees is large. When $p$
is large the treatment is more likely to succeed and the number of 
enrolees is small. Since $p=1$ and $p=0$ represent the deterministic
the corresponding variance is zero. The variance is right-skewed because
the support of the distribution corresponding to successful treatment
is larger than that of failure.

@@

In the rest of this work we derive the distribution of the number of 
enrollees needed
to observe either $s$ successes or $t$ failures. We refer to this distribution
as the Stopped Negative Binomial (SNB). Some of its characteristics are
summarized in Tab.~\ref{tab:snb}.
This paper derives these results
and explores properties of the distribution.
%The next section introduces our notation and basic results
%including the density of the distribution along with a description of
%its relation to other distributions. 
Section 2 derives the distribution function
based on a defined Bernoulli process and gives some basic properties.
Section 3 shows how the distribution is related to other standard
distributions and connects the SNB tail probability to the Binomial tail 
probability.
Section 4 derives the moment generating function.
Section 5 derives the compound distribution using a Beta distribution
for $p$.

@@

\begin{figure}[bp!]
\centering
\subfloat[The compound distribution's expected sample size.]{\includegraphics[width=0.5\textwidth]{bayesian-sample-expectation.pdf}}
\hfill
\subfloat[The compound distribution's variance.]{\includegraphics[width=0.5\textwidth]{bayesian-sample-variance.pdf}}
\caption{
The expected size and variance of the compound SNB with $s=2$, $t=12$
with $\mathbb{E}\ p = 0.045$ varying the uncertainty 
in the distribution of $p$. A larger value of $c$
corresponds to a Beta with smaller variance.
}
\label{fig:bayesian-sample-size}
\end{figure}
